---
title: "Include Odg Graphics"
author: "Peter von Rohr"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Include Odg Graphics}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r src_doc_info_setup}
this_rmd <- rprojroot::thisfile()
s_this_file <- "include_odg.Rmd"
this_rmd_fun <- rprojroot::is_rstudio_project$make_fix_file()
#rprojroot::find
this_rmd <- this_rmd_fun(s_this_file)
cat("This rmd: ", this_rmd, "\n")
```


## Disclaimer
This vignette describes the search for and the design for an easy to use functionality to include ODG-Graphics into an rmarkdown document. The ODG-Graphics file is not included directly into the rmd-document, but it is converted into a format such as pdf or png, that can then be used as graphics in an rmd-document.


## Introduction
One basic requirements for document authors is the ability to include graphics into a document. The technical details of how that is done depends on the document authoring system that is used. 

### WYSIWYG
Most such systems have their own graphics functionality included. Alternatively it is also possible to prepare the graphics in a specialized program, and then include the graphics via drag-and-drop into the document or some menu-based functionality for graphics inclusion can be used. 


### LaTeX
According to lshort there are some packages to produce graphics directly in the LaTeX-source document. Graphics can also be produced outside of LaTeX and can then be included into the document using the function `\includegraphics`. 


### HTML
In HTML documents, graphics are produced in a separate program and are included using the `<img src>`-tag.


## Methods
Depending on the format of output produced from an rmd-document, graphics can be included in one of two ways according to LaTeX or according to HTML. The R-package `knitr` provides the function `include_graphics()` which can be used for both output formats. That is why, we will be using this functionality.  

This package `rmddochelper` has already an existing method of creating an odg-graphic using the function `create_odg_graphic()`. This function is called not from within a document, but outside of the document in a separate R-console (e.g. in RStudio). This function is rather complicated to use, because it offers many parameters. A first approach might very well be to simplify this function. This is postponed to a later point in time, in favor of building a function for a smoother workflow which is described below.


## Workflow
An author is writing a document and decides to include a graphic into that document. To be able to do that, this graphic has to be created first. One possibility to do that is to run a function called `rmddochelper::use_odg_graphic` inside an R-code-chunk as shown below

```{r my_first_graphic, eval=FALSE}
rmddochelper::use_odg_graphic(ps_odg_file = "my_first_graphic.odg")
```

When executing this function inside the Rmd-source-file, an odg-file is created in the directory where this rmd-file is located. The above shown function is supposed to work like the `devtools::use_*` class of functions. An important pre-requisite of the shown function is that we have a tool available to create a graphic in an odg-format. One example of such a tool is `LibreOffice draw`. 




The question is how we can decide the location of this rmd-document. One possibility is to use the `rprojroot` package. The function `thisfile()` returns the absolute path to this rmd-document as shown below.

```{r}
this_rmd_rprojroot <- rprojroot::thisfile()
cat("This file is: ", this_rmd, "\n")
```

But the solution with `rprojroot` works only when the rmd-document is transformed using `knitr` because the determination of the path to the current rmd-source document is based on functionality in `knitr`. Hence when trying to run a chunk inside this document interactively as in a notebook, the solution using `rprojroot` does not work.



Depending on the parent directory of this rmd-document, we want to place the created odg-graphic in a different place. In case this rmd-source document is placed in a separate subdirectory called `rmd`, we put all odg-graphics under `../odg`. Otherwise, the odg-graphic is just put in a subdirectory called `odg`. 

```{r}
par_dir <- basename(dirname(this_rmd))
cat("Parent directory: ", par_dir, "\n")
# determine odg_dir
get_odg_dir4this_rmd <- function(ps_this_rmd){
  # determine parent directory of ps_this_file
  s_parent_dir <- dirname(ps_this_rmd)
  s_base_parent_dir <- basename(s_parent_dir)
  s_odg_parent_dir <- s_parent_dir
  if (s_base_parent_dir == "rmd") 
    s_odg_parent_dir <- dirname(s_parent_dir)
  return(file.path(s_odg_parent_dir, "odg"))
  
}
# output
cat("Current odg_dir: ", get_odg_dir4this_rmd(ps_this_rmd = this_rmd))
cat("Test odg_dir: ", get_odg_dir4this_rmd(ps_this_rmd = file.path(dirname(this_rmd), "rmd", basename(this_rmd))))
```

At this point we are ready to place the created odg-graphic in an automatically determined subdirectory. This subdirectory is determined using the function `get_odg_dir4this_rmd()`. The call to the cration function would then be

```{r my_second_odg, eval=FALSE}
s_odg_dir <- get_odg_dir4this_rmd(ps_this_rmd = this_rmd)
s_odg_file <- file.path(s_odg_dir, "my_second_odg.odg")
rmddochelper::use_odg_graphic(ps_odg_file = s_odg_file)
```




## List of Source Document Information


